---
title: "Analysing Bank Marketing Campaigns Using Logistic Regression and `tidymodels`"
subtitle: "Which factors drive a customer to open a Term Deposit with a bank?"
author: "John Karuitha"
date: "`r format(Sys.Date(), '%A, %B %d, %Y')`"
output:
  rmdformats::downcute:
    self_contained: true
    code_folding: hide
bibliography: citations.bib
csl: apa.csl
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(tidymodels)
library(data.table)
library(janitor)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## **Background**

In this article, I use data from a commercial bank to develop a model for predicting whether or not a customer will subscribe to a term deposit. The data comes from the [University of California at Irvine](http://archive.ics.uci.edu/ml/datasets/Bank+Marketing) database ^[Please follow http://archive.ics.uci.edu/ml/datasets/Bank+Marketing for a copy of the data], based originally on real data from a commercial bank in Portugal [@moro2014data]. Researchers have enriched the dataset to include additional macroeconomic variables from Banco de Portugal and publicly available at <https://www.bportugal.pt/estatisticasweb>. I have also loaded the data in my [Github repository](https://raw.githubusercontent.com/Karuitha/tidymodels_exe/master/bank-additional-full.csv). You can find the link to the data in my Github account in the footnotes ^[See, https://raw.githubusercontent.com/Karuitha/tidymodels_exe/master/bank-additional-full.csv]. 

When sales executives reach out to customers, the goal is to make a persuasive case so that the customer will purchase a product or products and hence improve the banks sales and profitability. In this case, the data concerns customers contacted by the bank's sales executives regarding a term deposit service. The aim of the exercise is to develop a model to predict, based on the given variables whether or not a customer will subscribe to a term deposit with the bank. 

## **Data and Variables**

The variables in the dataset fall into three categories; bank client data, data relating with the last contact with the customer in the current sales campaign, other attributes like number of contacts made with the customer, and country-level social and economic context attributes. I start by reading the data into `R`. 

```{r}
## I read in the data using data.table for speed
bank_data <- data.table::fread("https://raw.githubusercontent.com/Karuitha/tidymodels_exe/master/bank-additional-full.csv") %>% 
    
    ## Cleaning the column names to remove special characters and spaces
    janitor::clean_names()
```

The details of the variables are as follows; 

1. Age (numeric): Age in years of the customer. 

2. Job : type of job (categorical:"admin.", "blue-collar", "entrepreneur", "housemaid", "management", "retired", "self-employed", "services", "student", "technician", "unemployed", "unknown")

3. Marital : marital status (categorical: "divorced", "married", "single", "unknown"; note: "divorced" means divorced or widowed)

4. Education (categorical: "basic.4y", "basic.6y", "basic.9y", "high.school", "illiterate", "professional.course", "university.degree", "unknown"). 

5. Default: has credit in default? (categorical: "no","yes","unknown")

6. Housing: has housing loan? (categorical: "no","yes","unknown")

7. Loan: has personal loan? (categorical: "no","yes","unknown") related with the last contact of the current campaign:
    
8. Contact: contact communication type (categorical: "cellular","telephone")

9. Month: last contact month of year (categorical: "jan", "feb", "mar", …, "nov", "dec")

10. Dayofweek: last contact day of the week (categorical: "mon","tue","wed","thu","fri")

11. Duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y="no"). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.

12. Campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)

13. Pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)

14. Previous: number of contacts performed before this campaign and for this client (numeric)

15. Poutcome: outcome of the previous marketing campaign (categorical: "failure","nonexistent","success")

16. Emp.var.rate: employment variation rate - quarterly indicator (numeric)

17. Cons.price.idx: consumer price index - monthly indicator (numeric)

18. Cons.conf.idx: consumer confidence index - monthly indicator (numeric)

19. Euribor3m: euribor 3 month rate - daily indicator (numeric)

20. Nr.employed: number of employees - quarterly indicator (numeric)

21. Y: Output variable (desired target): Has the client subscribed a term deposit? (binary: "yes","no")

The dataset has `r nrow(bank_data)` observations for `r ncol(bank_data)` variables. The outcome variable is labelled `y` and is in the last column of the dataset. The other 20 variables are potential predictors. 

## **Train-Test Split**

I split the data into a training set and a testing set. I place 70% of the data into the training set while the balance goes to the testing set. I base the choice of the partition on my judgement, noting that we could use any other combination provided the bulk of the data goes to the training set. 

```{r}
## Create a split object
## Create a 70-30 split
bank_split <- bank_data %>% initial_split(prop = 0.7, strata = y)

## Training set
bank_train <- bank_split %>% training()

## Testing set
bank_test <- bank_split %>% testing()
```


## **Exploratory Data Analysis**

I start by reading in the data and performing exploratory data analysis. 

```{r}
## View the first six rows of the data
head(bank_data) %>% knitr::kable(caption = "First Six Rows of Bank Data")
    
```





## **References**

